Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
simplifying edits

\begin{definition}[Base induction variable]
  A variable \textit{v} of integer type is a \emph{base induction
    variable} if it is the control variable of a ``for''
  loop,\footnote{Or equivalent for the target language, \eg{},
    \texttt{do} in Fortran for our implementation} incremented by $1$ per
iteration. The variable is marked as a base induction variable
only within the scope of the loop. Induction variables are ranged
over by $i$, $j$, $k$ throughout. Since we do not use non-base
(derived) induction variables we often simply say \emph{induction variable}.
\end{definition}

@@

\begin{definition}[Neighbourhood index and subscript]
  For an array subscript $v(\bar{e})$, we say that $e \in \bar{e}$
  is a \emph{neighbourhood index} if it is of the form
  $e \equiv i$, $e \equiv i + a$ or $e \equiv i - a$ where $a$ is an
  integer constant. That is, a neighbourhood index is a
  relative index defined as a constant translation of a base induction
  variable. The relation $\equiv$ identifies terms up-to commutativity
  of $+$ and the inverse relation of $+$ and $-$ (\eg{},
  $(-b) + i \equiv i - b$).  We classify neighbourhood indices using
  the predicate \neigh{}. An array subscript $v(\bar{e})$ where every
 $e \in \bar{e}$ is a neighbourhood index is called a
 \emph{neighbourhood subscript}.
\label{def:neighbour}
\end{definition}

@@

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduces in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration). Specification declarations associate a
specification to one or more program variables $\bar{v}$ (here given
using the Fortran style syntax of $\ldots \; \texttt{::} \bar{v}$),
describing how the array variables $\bar{v}$ are accessed.
%Specification declarations are either \nonterm{spatial} or
%\nonterm{temporal} specifications and describe, for spatial
%specifications,
%Specifications declarations describe
%how an array variable $v$ is read
%, and for temporal
%specifications what variables are used to define the array variable $v$.
%%

@@

Regions have as terminals the \emph{region constants}, denoted by
\term{reflexive}, \term{forward}, \term{backward}, or \term{centered}. Each
region constant except \term{reflexive} is given a depth parameter $n$ (natural
number greater than 0). They all receive a dimension identifier $d$ (also a
natural number greater than 0). Region constants specify that an array is read
from by a collection of indices which in the $d^{th}$ dimension are
neighbourhood indices ranging from $i + 0$ up to $i \pm n$ inclusively,
\eg{}, the following is valid stencil, reading from \term{a} and writing to
\term{b}:
%%
\begin{minted}{fortran}
!= stencil forward(depth=2, dim=1) :: a
b(i, 0) = a(i, 0) + a(i+1, 0) + a(i+2, 0)
\end{minted}
%%
were \texttt{i} is an induction variable.  The stencil specification
is associated to the array variable \texttt{a}. Note that the second
dimension is fixed at a constant here.

A \term{backward} stencil
is similar but has negative neighbourhood indices, \eg{},
%
\begin{minted}{fortran}
!= stencil backward(depth=2, dim=1) :: a
b(i) = a(i) + a(i-1) + a(i-2)
\end{minted}
%
A \texttt{centered} stencil is a combination of \texttt{forward}
and \texttt{backward} stencils to the same depth, \eg{},
%%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1) :: a
b(i) = ( a(i) + a(i-1) + a(i+1) ) / 3.0
\end{minted}

>>>>

\begin{definition}[Base induction variable]
  A variable \textit{v} of integer type is a \emph{base induction
    variable} if it is the control variable of a ``for''
  loop,\footnote{Or equivalent in the source language, \eg{},
    \texttt{do} in Fortran for our implementation} incremented by $1$ per
iteration. The variable is marked as a base induction variable
only within the scope of the loop. Induction variables are ranged
over by $i$, $j$, $k$ throughout. Since we do not use non-base
(derived) induction variables we often simply say \emph{induction variable}.
\end{definition}

@@

\begin{definition}[Neighbourhood index]
  For an array subscript $v(\bar{e})$, we say that $e \in \bar{e}$
  is a \emph{neighbourhood index} if it is of the form
  $e \equiv i$, $e \equiv i + a$ or $e \equiv i - a$ where $a$ is an
  integer constant. That is, a neighbourhood index is a
  relative index defined as constant translations of base induction
  variables. (The relation $\equiv$ hre identifies terms up-to commutativity
  of $+$ and the inverse relation of $+$ and $-$ \eg{},
  $(-b) + i \equiv i - b$).  We classify neighbourhood indices using
  the predicate \neigh{}. %An array subscript $v(\bar{e})$ where every
% $e \in \bar{e}$ is a neighbourhood index is called a
% \emph{neighbourhood subscript}.
\label{def:neighbour}
\end{definition}

@@

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduce in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration). Specification declarations associate
(via the operator $\texttt{::}$) a specification to one or more program variables
$\bar{v}$, describing how the array variables $\bar{v}$ are accessed.
%Specification declarations are either \nonterm{spatial} or
%\nonterm{temporal} specifications and describe, for spatial
%specifications,
%Specifications declarations describe
%how an array variable $v$ is read
%, and for temporal
%specifications what variables are used to define the array variable $v$.
%%

@@

Regions have as terminals the \emph{region constants}, denoted by
\term{reflexive}, \term{forward}, \term{backward}, or \term{centered}. Each
region constant except \term{reflexive} is given a depth parameter $n$ (natural
number greater than 0). Each receives a dimension identifier $d$ (also a
natural number greater than 0). Region constants specify that an array is read
from by a collection of indices which in the $d^{th}$ dimension are
neighbourhood indices ranging from $i + 0$ up to $i \pm n$ inclusively,
\eg{}, the following is a valid stencil, reading from \term{a} and writing to
\term{b}:
%%
\begin{minted}{fortran}
!= stencil forward(depth=2, dim=1) :: a
b(i, 0) = a(i, 0) + a(i+1, 0) + a(i+2, 0)
\end{minted}
%%
were \texttt{i} is an induction variable.  The stencil specification
is associated to the array variable \texttt{a}. Note that the second
dimension is fixed at a constant here.
A \term{backward} stencil
is similar but has negative neighbourhood indices, \eg{},
%
\begin{minted}{fortran}
!= stencil backward(depth=2, dim=1) :: a
b(i) = a(i) + a(i-1) + a(i-2)
\end{minted}
%
A \texttt{centered} stencil is a combination of \texttt{forward}
and \texttt{backward} stencils to the same depth, \eg{},
%%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1) :: a
b(i) = ( a(i) + a(i-1) + a(i+1) ) / 3.0
\end{minted}
